---
title: "Data Storytelling"
author: "Grishma Bhattarai"
date: "4/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Divvy
## Exploratory Data Analysis

```{r}
# Import the data and look at the first six rows
df <- read.csv(file = 'divvy_trip_history_201909-202108.csv')
head(df)
```
### Over what period of time does the provided Divvy data cover?
```{r}
range(df$date)
```
Divvy data covers time period from 2019-09-01 to 2021-08-31.

### Considering only rides that started in 2021, which 5 stations are most used for ride starts? Please include the count of rides for each of these top stations in your results.
```{r}
library('tidyr')
library('dplyr')
```

```{r}
df %>% 
  filter(year == 2021)%>%
  count(start_station_name)%>%
  arrange(desc(n))
```
Top 5 stations most used for start ride:
1. NA (363123)
2. Streeter Dr & Grand Ave (61536)
3. Michigan Ave & Oak St (32968)
4. Millennium Park (29994)
5. Wells St & Concord Ln (29218)


### What about trip ends? Please include the count of rides for each of these top stations in your results.
```{r}
df %>% 
  filter(year == 2021)%>%
  count(end_station_name)%>%
  arrange(desc(n))
```
Top 5 stations most used for ride ends:
1. NA- No Station Name (386126)
2. Streeter Dr & Grand Ave (61953)
3. Michigan Ave & Oak St (33340)
4. Millennium Park (30525)
5. Wells St & Concord Ln (29496)

### What are the Top 5 trip start stations for members and top 5 trip start stations for casual riders?
```{r}
df %>% 
  filter(year == 2021)%>%
  group_by(member_casual)%>%
  count(start_station_name)%>%
  arrange(desc(n))
```
The top 5 trip start stations for members are: 
1. NA
2. Clark St & Elm St
3. Wells St & Concord Ln
4. Kingsbury St & Kinzie St
5. Wells St & Elm St
Similarly, the top 5 trip start stations for casual riders are:
1. NA
2. Streeter Dr & Grand Ave
3. Millennium Park
4. Michigan Ave & Oak St
5. Lake Shore Dr & Monroe St

### What was the mean ride duration in 2020? What was the median ride duration in 2020? 

```{r}
library(lubridate)
df$started_at <- as.POSIXct(df$started_at) 
df$ended_at <- as.POSIXct(df$ended_at) 
```

```{r}
head(df)
```

```{r}
df$interval <- difftime(df$ended_at, df$started_at, units='mins')
```

```{r}
df %>% 
  filter(year == 2020)%>%
  summarise(mean=mean(interval))
```
```{r}
df %>% 
  filter(year == 2020)%>%
  summarise(median=median(interval))
```


### What was the mean and median ride duration so far in 2021? 
```{r}
df %>% 
  filter(year == 2021)%>%
  summarise(mean2=mean(interval))
```

```{r}
df %>% 
  filter(year == 2021)%>%
  summarise(median2=median(interval))
```


### How many rides occurred in July 2021? 

```{r}
df %>% 
  filter(year == 2021, month == 7)%>%
  nrow
```

### How many rides occurred in January 2021?
```{r}
df %>% 
  filter(year == 2021, month == 1)%>%
  nrow
```

```{r}
df$weekday <- wday(df$date, label=TRUE)
```

```{r}
library(ggplot2)
```

```{r}
df %>%
  group_by(weekday, member_casual, year)%>%
  summarise(cnt=n())%>%
  ggplot(aes(fill=member_casual, y=cnt, x=weekday)) + 
  geom_bar(position="dodge", stat="identity") +
  facet_wrap(~year) + 
  theme_minimal() + scale_fill_manual(values = c("grey", "sky blue"), labels = c("Casual", "Member")) +
  scale_y_continuous(labels = scales::label_number()) +
  labs(title = "We Love the Weekend and So Do Divvy Riders",
       subtitle = "Demand for Divvy bikes fluctuates depending on the day of the week– the closer 
       the weekend, the higher the demand.",
       y = "Total Number of Rides",
       x = "Day of the Week",
       fill = 'Membership Status') + 
  theme(plot.subtitle = element_text(size = 8))
```



```{r}
df %>%
  mutate(date = ymd(date)) %>%
  ggplot(aes(x = date),fig(15,4)) +
  annotate("rect", xmin = ymd("2020-06-21"), xmax = ymd("2020-09-22"), ymin = -Inf, ymax = Inf, 
          fill="orange2", alpha=0.15) +
  annotate("rect", xmin = ymd("2019-09-01"), xmax = ymd("2019-09-22"), ymin = -Inf, ymax = Inf, 
          fill="orange2", alpha=0.15) +
    annotate("rect", xmin = ymd("2021-06-01"), xmax = ymd("2021-09-22"), ymin = -Inf, ymax = Inf, 
          fill="orange2", alpha=0.15) +
  annotate("rect", xmin = ymd("2020-12-21"), xmax = ymd("2021-03-15"), ymin = -Inf, ymax = Inf, 
          fill="light blue", alpha=0.18) +
  annotate("rect", xmin = ymd("2019-12-21"), xmax = ymd("2020-03-15"), ymin = -Inf, ymax = Inf, 
          fill="light blue", alpha=0.18) +
  geom_freqpoly(aes(color = member_casual), binwidth = 30, size=0.75) +
  scale_y_continuous(labels = scales::label_number()) + theme_classic()+
  scale_color_manual(values = c("gray35", "deepskyblue"), labels = c("Casual", "Member")) +
  labs(title = "Grim Weather = Grim Divvy Demand",
       subtitle = "Demand for Divvy bikes is seasonal– it fluctuates depending on whether it is summer or winter.",
       y = "Total Number of Rides",
       x = "Date",
       color = 'Membership Status') + 
  theme(plot.subtitle = element_text(size = 8))
```




```{r}

df %>%
  mutate(date = ymd(date)) %>%
  group_by(date)%>%
  summarise(cnt=n())%>%
  ggplot(aes(x = date, y=cnt),fig(15,4)) +
  geom_point(color='gray35') +
  stat_smooth(colour = 'deepskyblue') +
  geom_vline(xintercept = ymd("2020-03-15"), linetype = "dashed") +
  geom_vline(xintercept = ymd("2020-07-24"), linetype = "dashed") +
  geom_vline(xintercept = ymd("2021-01-19"), linetype = "dashed") +
  geom_vline(xintercept = ymd("2021-05-18"), linetype = "dashed") +
  scale_y_continuous(labels = scales::label_number()) + theme_classic() +
  labs(title = "Covid and Cycling",
       subtitle = "Demand for Divvy bikes fluctuates depending on the pandemic restrictions.",
       y = "Total Number of Rides",
       x = "Date") +
  geom_label(x=ymd("2020-03-15"),y=13000,label="Onset of Pandemic", size=2) +
  geom_label(x=ymd("2020-07-24"),y=32000,label="Gradual Resumption", size=2) +
  geom_label(x=ymd("2021-01-19"),y=11000,label="Stay-at-Home \n Advisory Re-Issued", size=2)+
  geom_label(x=ymd("2021-05-18"),y=14000,label="Bridge Phase", size=2)+ 
  theme(plot.subtitle = element_text(size = 8))

```


